\section{Feedforward Networks}
\emph{Deep feedforward networks}, also called \emph{feedforward neural networks} or \emph{multilayer perceptrons} (MLPs) are are basic deep learning models.
Their main purpose is to approximate some kind of function \(f^*(x)\).
Take for example a classification function \(y = f^*(x)\), where every input \(x\) maps to a class \(y\).
A feedforward network now aims to learn the best parameters \(\Theta\) for a function \(y = f(x, \Theta)\) that approximates \(f^*(x)\).

These models are \emph{networks} due to the fact that they can be represented by \emph{directed graphs}.
One of those \emph{graphs} models the consecutive execution of functions whose composition is \(f\).
They are called \emph{feedforward networks} because the input \(x\) flows through these functions without any \emph{feedback} connections.
This means that the graph is \emph{acyclic}.
If the graph is \emph{not acyclic} and therefore \(feedback\) connections exist, we speak of \emph{recurrent neural networks}.

For example let \(f(x) = (f^3 \circ f^2 \circ f^1)(x) = f^3(f^2(f^1(x)))\).
We now define the \emph{depth} of a \emph{feedforward network} \(f\) as the number of functions that compose it.
Our example has a \emph{depth} of \(3\).
These kind of chains are very typical in \emph{neural networks}, as each function in the chain represents a layer in the \emph{directed acyclic graph}.
The last element of the chain (in our example \(f^1\)) and therefore the last layer in the graph is called the \emph{output layer} as it maps the input \(x\) to its final value.
If we think of \(f\) as a classifier function the elements of the \emph{codomain} of \(f^1\) are the classes.
Likewise, the elements of the \emph{domain} of the first function in the chain (in our example \(f^3\)) are the inputs of the network.
The first layer is called \emph{input layer}.
Layers between the \emph{input} and \emph{output} layer are called \emph{hidden layers}.

\subsection{Feedforward network graphs}
As proposed earlier we can also think of \emph{neural networks} as \emph{acyclic directed graphs}.
Loosely inspired by neuroscience, \emph{neurons} are interpreted as \emph{nodes} and \emph{synapses} as \emph{edges}.
Each \emph{node} now receives input from an arbitrary amount of neurons in the previous layer and computes an output with its own activation function.
An \emph{edges} on the other hand has a scalar \emph{weight}.
These \emph{weights} are now what can be adjusted to changed the output of the neural network.

It is shown that in most feedforward networks a \emph{rectifier linear units} (reLU) (\(f(x) = \text{max}(0, x)\)) activation function works best. \cite{Nair-Hinton} \cite{inproceedings}
However, in certain cases other activation functions might perform better.
Such would be for example the \emph{swish} activation function (\(f(x) = x * \text{sigmoid}(x)\)), that tends to perform better than reLU on deeper models. \cite{DBLP:journals/corr/abs-1710-05941}

% Small sample neural network graph
\input{feedforward-networks/neural-network-example.tex}

This DNN \fref{fig:dnn-example} can now also be described by its \emph{weight matrices}. 

We can represent the weights of the edges between two layers as a \emph{weight matrices} \(\fat{W} \in \mathbb{R}^{m \times n}\).
The input can be represented as a vector \({x \in \mathbb{R}^n}\).
Calculating the output of a neural network can therefore be achieved by iterative \emph{matrix-vector multiplication} and component-wise usage of the activation function on each layer.

For \fref{fig:dnn-example} the weight matrices are
\begin{equation}
    \fat{W}^{(0)} =
    \begin{pmatrix}
        w_{2}^{(0)} & w_{4}^{(0)} \\
        w_{3}^{(0)} & w_{5}^{(0)}
    \end{pmatrix}, \quad
    \fat{W}^{(1)} = 
    \begin{pmatrix}
        w_{1}^{(1)} & w_{2}^{(1)}
    \end{pmatrix}.
\end{equation}
With the corresponding bias vectors
\begin{equation}
    b^{(0)} =
    \begin{pmatrix}
        w_{0}^{(0)} \\
        w_{1}^{(0)}
    \end{pmatrix}, \quad
    b^{(1)} =
    \begin{pmatrix}
        w_{0}^{(1)}
    \end{pmatrix}.
\end{equation}
In order to get more \enquote{flexible} we add a \emph{bias} node to each layer.
This bias node always has an input of \(1\) and describes the constant in a linear equation.

Let \(f(x)\) now be the activation function and
\begin{equation}
    x =
    \begin{pmatrix}
        x_0 \\
        x_1
    \end{pmatrix},
\end{equation}
the input vector.
The output vector for \fref{fig:dnn-example} can be calculated by evaluating the following equations.
\begin{equation}
    x_0 =
    f(W^{(0)}x + b^{0}) =
    f(
    \begin{pmatrix}
        w_{2}^{(0)} & w_{4}^{(0)} \\
        w_{3}^{(0)} & w_{5}^{(0)}
    \end{pmatrix}
    \begin{pmatrix}
        x_0 \\
        x_1å
    \end{pmatrix}
    +
    \begin{pmatrix}
        w_{0}^{(0)} \\
        w_{1}^{(0)}
    \end{pmatrix}
    )
\end{equation}
\begin{equation}
    y =
    f(W^{(1)}x_0 + b^{1}) =
    f(
    \begin{pmatrix}
        w_{1}^{(1)} & w_{2}^{(1)}
    \end{pmatrix}
    x_0
    +
    \begin{pmatrix}
        w_{0}^{(1)}
    \end{pmatrix}
    )
\end{equation}
Where the activation function \(f\) is used component-wise.